An Embedding Framework for Consistent Polyhedral Surrogates

### 2019-05-22 neurips-19/poly-embed.tex
Clarifying hinge example

\raf{Give it another shot, but focus on flexing our formalism.  What does hinge elicit?  How can we see that it embeds (twice) 0-1 loss from our definitions?  Then maybe one sentence for why logistic does not.}
To illustrate the idea of embeddings, we now look in detail at two key examples.
First, consider a binary classification problem on the outcomes $\{-1, +1\}$.
Both the hinge loss $L_{hinge}(u,y) := (1 - uy)_+$ and logistic loss $L_{logistic}(u,y) := \ln\left(1 + \exp(-yu)\right)$ are convex surrogates that are minimized in expectation by predicting the mode, as is 0-1 loss $\ell_{0-1}(r,y) := \Ind[r\neq y]$.
However, we say that hinge loss embeds twice the 0-1 loss, while logistic loss does not embed 0-1 loss.
First, taking $\R = \{-1,1\}$ and the embedding to be the identity, we can write 
\begin{center}
\begin{tabular}{ll}
	$\inprod{p}{\ell_{0-1}(1)} = (1-p_1)$ & $\inprod{p}{\ell_{0-1}(-1)} = (p_1)$\\
	$\inprod{p}{L_{hinge}(1)} = 2(1-p_1)$ & $\inprod{p}{L_{hinge}(-1)} = 2(p_1)$
\end{tabular}
\end{center}

Thus, for every $r \in \R$ and $p \in \simplex$, we have $2 \inprod{p}{\ell_{0-1}(r)} = \inprod{p}{L_{hinge}(\varphi(r))}$.

Now to see that optimal reports map to optimal reports, consider that 
\[
\prop{2 \ell_{0-1}} = \begin{cases}
1 & p_1 > 1/2 \\
\{-1,1\} & p_1 = 1/2\\
-1 & p_1 < 1/2
\end{cases}
\quad
\prop{L_{hinge}} = \begin{cases}
[1,\infty) & p_1 = 1\\
1 & p_1 \in (1/2,1) \\
[-1,1] & p_1 = 1/2\\
-1& p_1 \in (0, 1/2)\\
(-\infty, -1)& p_1 = 0
\end{cases}
\]

One can see logistic loss does not embed $\ell_{0-1}$ because it elicits a continuous property, so there is an infinite set of unique optima over $\simplex$, and thus no injection exists mapping this set to $\{-1,1\}$.

@@

  \raft{Jessie: I commented out your version below the proof.  What I have is on the informal side (so not necessarily better than what you had), mostly because of the ``working in the affine hull'' bit.  It can all be formalized though: just take an appropriate linear projection of $\reals^n$ to $\reals^{n-1}$, and work with the projected versions of everything, and then project back.  I can explain more on a call...}
  For the reverse implication, assume that $\risk{L} = \risk{\ell}$.
  In what follows, we implicitly work in the affine hull of $\simplex$, so that interiors are well-defined, and $\risk{\ell}$ may be differentiable on the (relative) interior of $\simplex$.
  Since $\ell$ is discrete, $-\risk{\ell}$ is polyhedral as the pointwise maximum of a finite set of linear functions.
  The projection of its epigraph $E_\ell$ onto $\simplex$ forms a power diagram by Theorem~\ref{thm:aurenhammer}, whose cells are full-dimensional and correspond to the level sets $\gamma_r$ of $\gamma = \prop{\ell}$.

@@

Now, for example, consider the surrogate discussed by Lapin et al.~\cite{yang2018consistency}, and Yang and Koyejo~\cite{yang2018consistency}.
We write the surrogate $L_2(u,y) := \max\left( \frac{1}{k} \sum_{i=1}^k (u + \indopp(y))_{[i]} - u_y \right) $, where $\indopp(y) = \ones - e_y$, and $(x)_{[i]}$ is the $i^{th}$ largest component of $x$.

Consider the following process:  (1.) For all $p \in \simplex$, we can show that there is some $u \in \{0,1\}^n$ such that $u \in \argmin_{u'} \inprod{p}{L_2(u')}$.
(2.) Now consider the discrete loss
\begin{align*}
\ell_2(r,y) &= 1 - \Ind[y \in r] + \min\left(\frac{|r| \setminus \{y\}}{k}, 1\right)
\end{align*}
We claim that $L_2$ embeds $\ell_2$.  
(3.) Finally, taking $k = 1$, we see that $0 \in \argmin_{r' \subseteq [n]} \inprod{p}{\ell_2(r')}$ for any $p$ such that $\max_y p_y \leq 1/2$.

>>>>

\raf{Give it another shot, but focus on flexing our formalism.  What does hinge elicit?  How can we see that it embeds (twice) 0-1 loss from our definitions?  Then maybe one sentence for why logistic does not.}
To illustrate the idea of embeddings, we now look in detail at two key examples.
First, consider a binary classification problem on the outcomes $\{-1, +1\}$.
Both the hinge loss $L_{hinge}(u,y) := (1 - uy)_+$ and logistic loss $L_{logistic}(u,y) := \ln\left(1 + \exp(-yu)\right)$ are convex surrogates that are minimized in expectation by predicting the mode, as is 0-1 loss $\ell_{0-1}(r,y) := \Ind[r\neq y]$.
However, we say that hinge loss embeds twice the 0-1 loss, while logistic loss does not embed 0-1 loss.

First, consider that 
\begin{align*}
	\ell_{0-1}(-1) = (0,1) && \ell_{0-1}(1) = (1,0)\\
	L_{hinge}(-1) = (0,2) && L_{hinge}(1) = (2,0)\\
\end{align*}
Therefore, $L_{hinge}$ is twice $\ell_{0-1}$ for each $r \in \R$.

Now to see that optimal reports map to optimal reports, consider that 
\[
\prop{\ell_{0-1}} = \begin{cases}
1 & p_1 > 1/2 \\
\{-1,1\} & p_1 = 1/2\\
-1 & p_1 < 1/2
\end{cases}
\qquad 
\prop{L_{hinge}} = \begin{cases}
[1,\infty) & p_1 = 1\\
1 & p_1 \in (1/2,1) \\
[-1,1] & p_1 = 1/2\\
-1& p_1 \in (0, 1/2)\\
(-\infty, -1]& p_1 = 0
\end{cases}
\]

Now, take the embedding $\varphi$ to be the identity.  % := r$ if $r \in [-1,1]$, and $\sgn(r)$ otherwise.
By inspection, we can see that if $r \in \prop{\ell_{0-1}}$, then $\varphi(r) = r \in \prop{L_{hinge}}$.  
Concretely, $-1 \in \prop{\ell_{0-1}}$ and $\prop{L_{hinge}}$ if and only if $p_1 \in [0, 1/2]$. 
Similarly, $1 \in \prop{\ell_{0-1}}$ and $\prop{L_{hinge}}$ if and only if $p_1 \in [1/2, 1]$.

Taking $\R = \{-1,1\}$ and the above embedding, we can write the expected loss over each report
\begin{center}
	\begin{tabular}{ll}
		$\inprod{p}{\ell_{0-1}(1)} = (1-p_1)$ & $\inprod{p}{\ell_{0-1}(-1)} = (p_1)$\\
		$\inprod{p}{L_{hinge}(1)} = 2(1-p_1)$ & $\inprod{p}{L_{hinge}(-1)} = 2(p_1)$
	\end{tabular}
\end{center}
Thus, for every $r \in \R$ and $p \in \simplex$, we have $2 \inprod{p}{\ell_{0-1}(r)} = \inprod{p}{L_{hinge}(\varphi(r))}$.
This leads us into the next section, where we show the Bayes Risks of the losses will be the same, accounting for the multiplicative constant, if one loss embeds the other.

Additionally, one can see logistic loss does not embed $\ell_{0-1}$ because it elicits a continuous property, so there is an infinite set of unique optima over $\simplex$, and thus no injection exists mapping this set to $\{-1,1\}$.

@@

  %\raft{Jessie: I commented out your version below the proof.  What I have is on the informal side (so not necessarily better than what you had), mostly because of the ``working in the affine hull'' bit.  It can all be formalized though: just take an appropriate linear projection of $\reals^n$ to $\reals^{n-1}$, and work with the projected versions of everything, and then project back.  I can explain more on a call...}
  For the reverse implication, assume that $\risk{L} = \risk{\ell}$.
  In what follows, we implicitly work in the affine hull of $\simplex$, so that interiors are well-defined, and $\risk{\ell}$ may be differentiable on the (relative) interior of $\simplex$.
  Since $\ell$ is discrete, $-\risk{\ell}$ is polyhedral as the pointwise maximum of a finite set of linear functions.
  The projection of its epigraph $E_\ell$ onto $\simplex$ forms a power diagram by Theorem~\ref{thm:aurenhammer}, whose cells are full-dimensional and correspond to the level sets $\gamma_r$ of $\gamma = \prop{\ell}$.

@@

Now, for example, consider the surrogate discussed by Lapin et al.~\cite{yang2018consistency}, and Yang and Koyejo~\cite{yang2018consistency}.
We write the surrogate $L_2(u,y) := \max\left( \frac{1}{k} \sum_{i=1}^k (u + \indopp(y))_{[i]} - u_y \right)_+ $, where $\indopp(y) = \ones - e_y$, and $(x)_{[i]}$ is the $i^{th}$ largest component of $x$.

Consider the following process:  (1.) For all $p \in \simplex$, we can show that there is some $u \in \{0,1\}^n$ such that $u \in \argmin_{u'} \inprod{p}{L_2(u')}$.
(2.) Now consider the discrete loss
\begin{align*}
\ell_2(r,y) &= \Ind[y \not\in r] + \min\left(\frac{|r \setminus \{y\}|}{k}, 1\right)
\end{align*}
We claim that $L_2$ embeds $\ell_2$.  
(3.) Finally, taking $k = 1$, we see that $0 \in \argmin_{r' \subseteq [n]} \inprod{p}{\ell_2(r')}$ for any $p$ such that $\max_y p_y \leq 1/2$.
